# Assignment for Week 2 - Reproducible Research
### Author: APS

This is an assignment for peer assessment for Reproducible Research.

## Loading the data
First we will unzip the data and read the file.

```{r, cache=TRUE, echo=TRUE}
unzip("activity.zip")
activity <- read.csv("./activity.csv", sep = ",")
str(activity)
head(activity)
```

## Processing the data

Here, I convert date column to Date class from factor.

```{r, echo=TRUE}
activity$date <- as.Date(activity$date, "%Y-%m-%d")
```

## Mean total number of steps taken per day

1. Here is the code to calculate the total number of steps taken each day, and plot this as a histogram.

```{r histogram, echo=TRUE}
activity1 <- aggregate(steps ~ date, activity, sum)
head(activity1)
hist(activity1$steps, xlab = "Total number of Steps taken each day", main = "Histogram of total steps")
```

2. Here is the code to calculate mean number of steps taken per day.

```{r activiy.mean, echo=TRUE}
activity.mean <- aggregate(steps ~ date, activity, mean)
activity.mean
```

Here is the code to calculate median number of steps taken per day.

```{r activity.median, echo=TRUE}
activity.median <- aggregate(steps ~ date, activity, median)
activity.median
```

## Average daily activity pattern

1. Calculate daily activity pattern.

```{r activityPattern, echo=TRUE}
activity.int <- aggregate(steps ~ interval, activity, mean)
with(activity.int, plot(interval, steps, type = "l", main = "Time-series plot of average daily activity pattern"))
```

2. Code to calculate the 5-minute interval that contains maximum number of steps on average across all the days in the dataset.

```{r maximumActivity, echo=TRUE}
max.int <- activity.int[which.max(activity.int$steps), ]
x <- max.int$interval
```

Interval `r x` contains maximum number of steps n average across all the days in the dataset.

## Imputing missing values

1. Calculate the total number of missing value in the dataset.

```{r missing.values, echo=TRUE}
na.data <- sum(is.na(activity))
```

Data contains `r na.data` missing values.

2. Fill missing values with the mean value for the 5-minute interval

This code imputes missing values with the mean for that 5-minute interval.

```{r fill.missing.values, echo=TRUE}
library(zoo)
impute.activity <- activity
impute.activity$steps <- na.aggregate(impute.activity$steps, by = impute.activity$interval)
head(impute.activity)
```

3. With the abovementioned code, *impute.activity* - a new data set, is created that is equal to the original dataset with the missing data filled in.

4. This code generates a histogram for the imputed dataset.

```{r histogram.imputed.data, echo=TRUE}
impute.activity1 <- aggregate(steps ~ date, impute.activity, sum)
hist(impute.activity1$steps, xlab = "Total number of Steps taken each day", main = "Histogram of total steps")
```

Following code calculates mean number of steps taken per day.

```{r mean.imputed.activity, echo=TRUE}
imputed.mean <- aggregate(steps ~ date, impute.activity, mean)
imputed.mean
```

Following code calculates median number of steps taken per day.

```{r median.imputed.activity, echo=TRUE}
imputed.median <- aggregate(steps ~ date, impute.activity, median)
imputed.median
```

Imputed values don't change the general pattern of the histogram. However, median values are significantly affected.

## Differences in activity patterns between weekdays and weekends

1. Create a factor variable in the dataset with two levels -- "weekday" and "weekend".

```{r weekdays, echo=TRUE}
library(timeDate)
library(dplyr)
week.activity <- activity %>% mutate(weekday = isWeekday(activity$date))
head(week.activity)
week.activity <- transform(week.activity, weekday = factor(weekday))
```

The code above creates a logical factor variable indicating weekdays (TRUE) and weekends (FALSE).

2. Here is the code to generate lattice plot showing differences in activity patterns between weekdays and weekends.

```{r plot, echo=TRUE}
library(lattice)
xyplot(steps ~ interval | weekday, data = week.activity, layout = c(1, 2), type = "l", strip = strip.custom(factor.levels = c("Weekday", "Weekend")))
```

