# Assignment for Week 2 - Reproducible Research
### Author: APS

This is an assignment for peer assessment for Reproducible Research.
First we will unzip the data and read the file.

```{r, cache=TRUE, echo=TRUE}
unzip("activity.zip")
activity <- read.csv("./activity.csv", sep = ",")
str(activity)
head(activity)
```

Here, I convert date column to Date class from factor.

```{r, echo=TRUE}
activity$date <- as.Date(activity$date, "%Y-%m-%d")
```

Here is the code to calculate the total number of steps taken each day, and plot this as a histogram.

```{r histogram, echo=TRUE}
activity1 <- aggregate(steps ~ date, activity, sum)
head(activity1)
hist(activity1$steps, xlab = "Total number of Steps taken each day", main = "Histogram of total steps")
```

Here is the code to calculate mean number of steps taken per day.

```{r activiy.mean, echo=TRUE}
activity.mean <- aggregate(steps ~ date, activity, mean)
activity.mean
```

Here is the code to calculate median number of steps taken per day.

```{r activity.median, echo=TRUE}
activity.median <- aggregate(steps ~ date, activity, median)
activity.median
```
